---
title: "REMA examples"
format: html
editor: visual
---

```{r}
#| echo: false
#| output: false

# install.packages("devtools")
# devtools::install_github("afsc-assessments/rema", dependencies = TRUE, build_vignettes = FALSE)
library(rema)
libs <- c('readr', 'dplyr', 'tidyr', 'ggplot2', 'cowplot')
if(length(libs[which(libs %in% rownames(installed.packages()) == FALSE )]) > 0) {install.packages(libs[which(libs %in% rownames(installed.packages()) == FALSE)])}
lapply(libs, library, character.only = TRUE)
ggplot2::theme_set(cowplot::theme_cowplot(font_size = 8) +
                     cowplot::background_grid() +
                     cowplot::panel_border())
```

## Simple case: Aleutian Islands Pacific cod

-   One survey index: AI bottom trawl biomass

-   Single time series (one area/stratum)

-   One fixed effect parameter estimated (process error SD)

```{r}
biomass_dat <- read_csv('ai_pcod_2022_biomass_dat.csv')

input <- prepare_rema_input(model_name = '2022 AI Pacific cod',
                            biomass_dat = biomass_dat)

m <- fit_rema(input)
```

You can add options to executable code like this

```{r}
out <- tidy_rema(m)
plot_rema(out)$biomass_by_strata
```

## Complex case: Gulf of Alaska Thornyheads

-   Two survey indices (GOA bottom trawl biomass and longline survey relative population weights; RPWs)

-   Trawl survey stratified by depth and area (9 strata)

    -   Western GOA, Central GOA, Eastern GOA x 0-500 m, 501-700 m, 701-1000 m

-   Longline survey stratified by area only (3 strata)

    -   Western GOA, Central GOA, Eastern GOA

-   Six fixed effects parameters

    -   Three process error SDs estimated (shared within each area across depth strata)

    -   One scaling parameter for the longline survey

    -   One additional observation error SD for the biomass survey (shared across all strata)

    -   One additional observation error SD for the longline survey (shared across all strata)

    ```{r}

    biomass_dat <- read_csv('goa_thornyhead_2022_biomass_dat.csv') # trawl biomass
    cpue_dat <- read_csv('goa_thornyhead_2022_cpue_dat.csv') # longline RPWs

    input <- prepare_rema_input(model_name = '2022 GOA thornyheads',
                                multi_survey = TRUE,
                                biomass_dat = biomass_dat,
                                cpue_dat = cpue_dat,
                                # RPWs are a summable/area-weighted effort index
                                sum_cpue_index = TRUE, 
                                # three process error parameters (log_PE) estimated,
                                # indexed as follows for each biomass survey stratum
                                # (shared within an area across depths):
                                PE_options = list(pointer_PE_biomass = c(1, 1, 1, 2, 2, 2, 3, 3, 3)),
                                # scaling parameter options:
                                q_options = list(
                                  # longline survey strata (n=3) indexed as follows for the
                                  # biomass strata (n=9)
                                  pointer_biomass_cpue_strata = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
                                  # one scaling parameters (log_q) estimated, shared
                                  # over all three LLS strata
                                  pointer_q_cpue = c(1, 1, 1)),
                                # estimate extra trawl survey observation error
                                extra_biomass_cv = list(assumption = 'extra_cv'),
                                # estimate extra longline survey observation error
                                extra_cpue_cv = list(assumption = 'extra_cv'))

    m <- fit_rema(input)
    ```

```{r}
out <- tidy_rema(m)
plots <- plot_rema(out)

cowplot::plot_grid(plots$biomass_by_strata +
                     theme(legend.position = 'none') +
                     geom_line() +
                     labs(x = NULL, y = NULL, subtitle = 'Trawl survey biomass (t)',
                          fill = NULL, colour = NULL, shape = NULL, lty = NULL),
                   plots$cpue_by_strata  +
                     facet_wrap(~strata, ncol = 1)  +
                     geom_line() +
                     labs(x = NULL, y = NULL, subtitle = 'Longline survey RPW',
                          fill = NULL, colour = NULL, shape = NULL, lty = NULL),
                   ncol = 2,
                   rel_widths = c(1.7, 1))
```

Plot the additional estimated observation error (bold error bar w/ whiskers = assumed CV from design-based survey estimates, full error bar w/ no whiskers = total (assumed + estimated):

```{r}
out <- tidy_extra_cv(out)
plots <- plot_extra_cv(out)

cowplot::plot_grid(plots$biomass_by_strata +
                     theme(legend.position = 'none') +
                     geom_line() +
                     labs(x = NULL, y = NULL, subtitle = 'Trawl survey biomass (t)',
                          fill = NULL, colour = NULL, shape = NULL, lty = NULL),
                   plots$cpue_by_strata  +
                     facet_wrap(~strata, ncol = 1)  +
                     geom_line() +
                     labs(x = NULL, y = NULL, subtitle = 'Longline survey RPW',
                          fill = NULL, colour = NULL, shape = NULL, lty = NULL),
                   ncol = 2,
                   rel_widths = c(1.7, 1))
```

## 
